\\fontsize{12pt}{3pt}\\selectfont \\textbf{\\textit{Submitted to the journal Ecological Monographs. Manuscript type: Article}}

### 2023-05-22 manuscript/manuscriptRW.Rmd
updates from EcolMod prep

---
title: "Tree regeneration in models of forest dynamics: a key priority for further research"
subtitle: "Submitted to the journal Ecological Monographs. Manuscript type: Article"
author: 
- Olalla Díaz-Yáñez$^{*1}$, 
- Yannek Käber$^1$, 
- Tim Anders$^2$,
- Friedrich Bohn$^3$,
- Kristin H. Braziunas$^4$,
- Josef	Brůna$^5$,
- Rico Fischer$^3$,
- Samuel M. Fischer$^3$,
- Jessica Hetzer$^2$,
- Thomas Hickler$^2$,
- Christian	Hochauer$^6$,
- Manfred J. Lexer$^6$,
- Heike	Lischke$^7$,
- Mats Mahnken$^8$,
- Paola	Mairota$^9$,
- Ján Merganič$^{10}$,
- Katarina Merganičová$^{11,12}$,
- Tobias Mette$^{13}$,
- Marco Mina$^{14}$,
- Xavier Morin$^{15}$
- Werner Rammer$^4$,
- Christopher P.O. Reyer$^8$,
- Simon	Scheiter$^2$,
- Daniel Scherrer$^{16}$,
- Harald Bugmann$^1$
date: $^1$ETH Zurich. Department of Environmental Systems Science. Institute of Terrestrial Ecosystems. Universitätstrasse 16, 8092 Zürich, Switzerland. \newline
      $^2$Senckenberg Biodiversity and Climate Research Centre (SBiK-F), 60325 Frankfurt am Main, Germany. \newline
      $^3$Department of Computational Hydrosystem, Helmholtz-Centre for Environmental Research - UFZ, Permoserstraße 15, 04318 Leipzig, Germany. \newline
      $^4$Ecosystem Dynamics and Forest Management Group, TUM School of Life Sciences, Technical University of Munich, 85354 Freising, Germany. \newline
       $^5$Department of Geoecology, Institute of Botany of the Czech Academy of Sciences, Průhonice, Czech Republic. \newline
       $^6$Institute of Silviculture, Department of Forest and Soil Sciences, University of Natural Resources and Life Sciences, Peter Jordan Strasse 82, A-1190 Vienna, Austria \newline
       $^7$Dynamic Macroecology, Swiss Federal Institute for Forest, Snow and Landscape Research WSL, Zürcherstrasse 111, 8903 Birmensdorf, Switzerland. \newline
       $^8$Potsdam Institute for Climate Impact Research (PIK), Member of the Leibniz Association, 14473 Potsdam, Germany. \newline
       $^9$Department of Soil Sciences, of Plants and Food, University of Bari Aldo Moro, Italy. \newline
       $^{10}$Department of Forest Harvesting, Logistics and Ameliorations, Faculty of Forestry, Technical University in Zvolen \newline
       $^{11}$Czech University of Life Sciences Prague, Faculty of Forestry and Wood Sciences, Kamýcká 129, 16500 Praha 6 – Suchdol, Czech Republic. \newline
       $^{12}$Department of Biodiversity of Ecosystems and Landscape, Slovak Academy of Sciences, Štefánikova 3, P.O.Box 25, 81499 Bratislava, Slovakia. \newline 
       $^{13}$Dpt. of Soil and Climate, Bavarian State Institute of Forestry. \newline 
       $^{14}$Institute for Alpine Environment, Eurac Research, Bolzano/Bozen, Italy. \newline 
       $^{15}$CEFE, CNRS Campus du CNRS 1919, route de Mende 34293 Montpellier cedex, France. \newline 
       $^{16}$Swiss Federal Institute for Forest, Snow and Landscape Research WSL, Zürcherstrasse 111, 8903 Birmensdorf, Switzerland.  \newline
       $^*$Corresponding author (olalla.diaz@usys.ethz.ch)
output:
  bookdown::pdf_document2:
    keep_tex: yes
    number_sections: yes
    toc: no
    extra_dependencies: "subfig"
  word_document:
    reference_docx: template_paper.docx
mainfont: Arial
fontsize: 12pt
linestretch: 1.5
bibliography: Regeneration.bib
link-citations: yes
nocite: |
  @bugmann1996a, @bohn2014, @fabrika2005, @mette2014, @smith2001a, @langan2017, @seidl2012a, @lischke2006, @schumacher2006, @scheiter2009, @scheller2007, @morin2021, @lasch-born2020, @hickler2012, @fischer_lessons_2016, @langan2017, @scheiter2013 
linkcolor: blue
csl: global-change-biology.csl
editor_options:
  markdown:
    wrap: 72
header-includes:
- \usepackage[left]{lineno}
- \linenumbers
- \usepackage{caption}
- \captionsetup[figure]{font=scriptsize}
- \captionsetup[table]{font=scriptsize}
- \usepackage{titling}
- \predate{\begin{flushleft}}
- \postdate{\end{flushleft}}
---

@@

OPEN RESEARCH STATEMENT: This submission uses novel code, both the data and code are provided in an external repository to be evaluated during the peer review process. The repository can be downloaded from [here](https://polybox.ethz.ch/index.php/s/q5MOB1CLcjARs3H) (password: DownDataCodeRegen1)]. The protocol is available at @díaz-yáñez2022. We will make publicly available the code and data after the revision.

**Keywords:** tree regeneration | tree recruitment | tree establishment | forest dynamics | models | landscape | stand | global

@@

```{r map, echo=FALSE, fig.cap="The 200 sites of the EuFoRIa network (EuFoRIa, 2019) used in the present study, the colors show the dominant tree species per site in the observed regeneration at the 10 cm threshold.", out.width='80%'}
knitr::include_graphics(here::here("figures", "theMap.png"))
```

@@

```{r over, echo=FALSE, fig.cap="Mean regeneration levels across all samples per site, plotted for the 200 sites and for each model. The red dashed lines shows the 25th and 75th percentiles for the 7 cm diameter threshold in the observed data. There are two boxplots for each model where the left lighter boxplot corresponds to 7 cm and the right darker boxplot to 10 cm", out.width='100%'}
        knitr::include_graphics(here::here("figures", "recOverAll710.png"))
```

@@

```{r H710, echo=FALSE, fig.cap="Shannon index of tree regeneration, calculated by basal area, as the mean value across the 200 samples per site for the diameter thresholds of 7 and 10 cm, respectively. The red dashed lines shows the 25th and 75th percentiles for the 7 cm diameter threshold in the observed data.", out.width='100%'}
    knitr::include_graphics(here::here("figures", "H7_10.png"))
```

@@

```{r combinedRichness, echo=FALSE, fig.cap="Mean Shannon index across all samples per site for observed and simulated data. Each plot shows one pattern represented by one exemplary model of each category (overperdicted, intermediate and underpredicted). The full data with the grouping of the models are shown in Figures S3 and S4. n indicates the number of models falling in each group. A) Examples for the three trends across models for species diversity at the stand level. B) Examples for the three trends across models for regeneration at the 7 cm threshold.", out.width='75%'}
knitr::include_graphics(here::here("figures", "combined_richness.png"))
```

@@

```{r mortality11, echo=FALSE, fig.cap="Ratio of tree regeneration rates between the 7 and 10 cm thresholds. Dashed blue lines mark a ratio equal to 1, indicating no decrease in tree regeneration between 7 and 10 cm, and a ratio equal to 1.77, corresponding to the Reineke self-thinning ratio under even-aged conditions.", out.width='75%'}
knitr::include_graphics(here::here("figures", "ratio7_10.png"))
```

@@

```{r mortality13, echo=FALSE, fig.cap="Ratio of tree regeneration between the 7 and 10 cm thresholds (regeneration 7 cm / regeneration 10 cm) and the overestimation proportion at 7 cm ([simulated - observed] / observed) for the mean regeneration per model across sites and samples. The horizontal dashed lines show a ratio equal to 1, indicating no decrease of regeneration between 7 and 10 cm, and a ratio equal to 1.77 corresponding to the Reineke self-thinning line under even-aged conditions. The vertical line indicates zero overestimation.", out.width='80%'}
knitr::include_graphics(here::here("figures", "mort7_10_medianOver.png"))
```

@@

```{r nvTrend710BAlightTemp, echo=FALSE, fig.cap="Mean regeneration values across the 200 samples per site with the y-axis scaled differently by model, for the 200 sites against gradients of (a) total basal area; (b) climatic water balance; (c) seasonal degree-day sum. The values were split into ten bins; the red lines represent a GAM model showing the trend in the observed data. For comparing the magnitude of recruitment across models the same graphs with equally scaled y-axis can be found in Figure S7.", fig.show = 'hold', out.width="49%",fig.ncol = 2, fig.subcap=c('', '', ''), fig.show='hold', fig.align='left'}

@@

```{r diff, echo=FALSE, fig.cap="Regeneration levels of the five main species in terms of their basal area share in the observations (top row) and the performance of each model across the environmental gradients (other rows) for the 7 cm regeneration threshold. The values shown are the mean of the 200 samples per site and across the sites in each bin (tile), with ten bins per gradient. The sizes of the circles represent the ratio between the regeneration basal area of the species and the total regeneration basal area for all species. The absence of a circle indicates a zero basal area share in the regeneration, or the absence of regeneration altogether. The color gradient (for the models only) shows the difference between the simulated and observed ratio of regeneration basal area of the species and the total regeneration basal area for all the species.", out.width='100%', out.height= '85%'}
knitr::include_graphics(here::here("figures", "HeatEnvGrad_10circles_7.png"))
```

@@

There were pronounced differences how the main tree species were represented by the models along the environmental gradients, in particular the dominant species *Abies alba*, *Fagus sylvatica* and *Picea abies*. The regeneration levels were sampled from simulations in the equilibrium, and in this state it is expected that non-climax species such as *Pinus sylvestris* or *Quercus* would be of minor importance, or absent [@klopcic2015]. Most models captured this low abundance, which is also found in the empirical data. Thus, although the broad patterns are matched by many models, improvements in the quantification of the regeneration niche of the species are needed, but this cannot be done in the absence of robust datasets across multiple stand development stages.

@@

# Author contributions
O. Díaz-Yáñez and Y. Käber contributed equally and share the first authorship. H. Bugmann, Y. Käber, and O. Díaz-Yáñez developed the study design. All authors contributed to preparing the simulations with their respective models. O. Díaz-Yáñez led the data analysis with contributions during the workshop from Y.Käber, T. Anders, K. H. Braziunas, J. Brůna, S. M. Fischer, J. Hetzer, T. Hickler, H. Lischke, M.Mahnken, P. Mairota, K. Merganičová, T. Mette, X. Morin, W. Rammer and D. Scherrer
H. Bugmann. O. Díaz-Yáñez and H. Bugmann led the writing of the manuscript. All authors participated in the revision of the manuscript and approved its submission.

@@

```{=html}
```

>>>>

---
subtitle: "\\fontsize{16pt}{3pt}\\selectfont \\textbf{\\textit{Tree regeneration in models of forest dynamics: a key priority for further research}} "
title: "\\fontsize{12pt}{3pt}\\selectfont \\textbf{\\textit{Submitted to the journal Ecological Monographs. Manuscript type: Article}}"
author: 
- Olalla Díaz-Yáñez$^{*1}$, 
- Yannek Käber$^1$, 
- Tim Anders$^2$,
- Friedrich Bohn$^3$,
- Kristin H. Braziunas$^4$,
- Josef	Brůna$^5$,
- Rico Fischer$^3$,
- Samuel M. Fischer$^3$,
- Jessica Hetzer$^2$,
- Thomas Hickler$^2$,
- Christian	Hochauer$^6$,
- Manfred J. Lexer$^6$,
- Heike	Lischke$^7$,
- Mats Mahnken$^8$,
- Paola	Mairota$^9$,
- Ján Merganič$^{10}$,
- Katarina Merganičová$^{11,12}$,
- Tobias Mette$^{13}$,
- Marco Mina$^{14}$,
- Xavier Morin$^{15}$
- Werner Rammer$^4$,
- Christopher P.O. Reyer$^8$,
- Simon	Scheiter$^2$,
- Daniel Scherrer$^{16}$,
- Harald Bugmann$^1$
date: $^1$ETH Zurich. Department of Environmental Systems Science. Institute of Terrestrial Ecosystems. Universitätstrasse 16, 8092 Zürich, Switzerland. \newline
      $^2$Senckenberg Biodiversity and Climate Research Centre (SBiK-F), 60325 Frankfurt am Main, Germany. \newline
      $^3$Department of Computational Hydrosystem, Helmholtz-Centre for Environmental Research - UFZ, Permoserstraße 15, 04318 Leipzig, Germany. \newline
      $^4$Ecosystem Dynamics and Forest Management Group, TUM School of Life Sciences, Technical University of Munich, 85354 Freising, Germany. \newline
       $^5$Department of Geoecology, Institute of Botany of the Czech Academy of Sciences, Průhonice, Czech Republic. \newline
       $^6$Institute of Silviculture, Department of Forest and Soil Sciences, University of Natural Resources and Life Sciences, Peter Jordan Strasse 82, A-1190 Vienna, Austria \newline
       $^7$Dynamic Macroecology, Swiss Federal Institute for Forest, Snow and Landscape Research WSL, Zürcherstrasse 111, 8903 Birmensdorf, Switzerland. \newline
       $^8$Potsdam Institute for Climate Impact Research (PIK), Member of the Leibniz Association, 14473 Potsdam, Germany. \newline
       $^9$Department of Soil Sciences, of Plants and Food, University of Bari Aldo Moro, Italy. \newline
       $^{10}$Department of Forest Harvesting, Logistics and Ameliorations, Faculty of Forestry, Technical University in Zvolen \newline
       $^{11}$Czech University of Life Sciences Prague, Faculty of Forestry and Wood Sciences, Kamýcká 129, 16500 Praha 6 – Suchdol, Czech Republic. \newline
       $^{12}$Department of Biodiversity of Ecosystems and Landscape, Slovak Academy of Sciences, Štefánikova 3, P.O.Box 25, 81499 Bratislava, Slovakia. \newline 
       $^{13}$Dpt. of Soil and Climate, Bavarian State Institute of Forestry. \newline 
       $^{14}$Institute for Alpine Environment, Eurac Research, Bolzano/Bozen, Italy. \newline 
       $^{15}$CEFE, CNRS Campus du CNRS 1919, route de Mende 34293 Montpellier cedex, France. \newline 
       $^{16}$Swiss Federal Institute for Forest, Snow and Landscape Research WSL, Zürcherstrasse 111, 8903 Birmensdorf, Switzerland.  \newline
       $^*$Corresponding author (olalla.diaz@usys.ethz.ch)
output:
  bookdown::pdf_document2:
    keep_tex: yes
    number_sections: yes
    toc: no
    extra_dependencies: "subfig"
    latex_engine: xelatex
  word_document:
    reference_docx: template_paper.docxl
fontsize: 12pt
linestretch: 2
bibliography: Regeneration.bib
link-citations: yes
nocite: |
  @bugmann1996a, @bohn2014, @fabrika2005, @mette2014, @smith2001a, @langan2017, @seidl2012a, @lischke2006, @schumacher2006, @scheiter2009, @scheller2007, @morin2021, @lasch-born2020, @hickler2012, @fischer_lessons_2016, @langan2017, @scheiter2013 
linkcolor: blue
csl: chicago-author-date.csl
editor_options:
  markdown:
    wrap: 72
header-includes:
- \usepackage[left]{lineno}
- \linenumbers
- \usepackage{caption}
- \captionsetup[figure]{font=scriptsize}
- \captionsetup[table]{font=scriptsize}
- \usepackage{titling}
- \predate{ }
- \postdate{  }
- \usepackage{fontspec}
- \setmainfont{Times New Roman}
---

@@

OPEN RESEARCH STATEMENT: This submission uses novel code, both the data
and code are provided in an external repository to be evaluated during
the peer review process. The repository can be downloaded from
[here](https://polybox.ethz.ch/index.php/s/q5MOB1CLcjARs3H) (password:
DownDataCodeRegen1). The protocol is available at @díaz-yáñez2022. We
will make publicly available the code and data after the revision.

**Keywords:** tree regeneration \| tree recruitment \| tree
establishment \| forest dynamics \| models \| landscape \| stand \|
global

@@

@@

@@

@@

@@

@@

@@

@@

@@

There were pronounced differences how the main tree species were
represented by the models along the environmental gradients, in
particular the dominant species *Abies alba*, *Fagus sylvatica* and
*Picea abies*. The regeneration levels were sampled from simulations in
the equilibrium, and in this state it is expected that non-climax
species such as *Pinus sylvestris* or *Quercus* would be of minor
importance, or absent [@klopcic2015]. Most models captured this low
abundance, which is also found in the empirical data. Thus, although the
broad patterns are matched by many models, improvements in the
quantification of the regeneration niche of the species are needed, but
this cannot be done in the absence of robust datasets across multiple

@@

# Author contributions

O. Díaz-Yáñez and Y. Käber contributed equally and share the first
authorship. H. Bugmann, Y. Käber, and O. Díaz-Yáñez developed the study
design. All authors contributed to preparing the simulations with their
respective models. O. Díaz-Yáñez led the data analysis with
contributions during the workshop from Y.Käber, T. Anders, K. H.
Braziunas, J. Brůna, S. M. Fischer, J. Hetzer, T. Hickler, H. Lischke,
M.Mahnken, P. Mairota, K. Merganičová, T. Mette, X. Morin, W. Rammer and
D. Scherrer H. Bugmann. O. Díaz-Yáñez and H. Bugmann led the writing of
the manuscript. All authors participated in the revision of the
manuscript and approved its submission.

@@

```{=html}
```
::: {#refs}
:::

# Table

# Figure captions

**Figure 1.** The 200 sites of the EuFoRIa network (EuFoRIa, 2019) used
in the present study, the colors show the dominant tree species per site
in the observed regeneration at the 10 cm threshold.

**Figure 2.** Mean regeneration levels across all samples per site,
plotted for the 200 sites and for each model. The red dashed lines shows
the 25th and 75th percentiles for the 7 cm diameter threshold in the
observed data. There are two boxplots for each model where the left
lighter boxplot corresponds to 7 cm and the right darker boxplot to 10
cm

**Figure 3.** Shannon index of tree regeneration, calculated by basal
area, as the mean value across the 200 samples per site for the diameter
thresholds of 7 and 10 cm, respectively. The red dashed lines shows the
25th and 75th percentiles for the 7 cm diameter threshold in the
observed data.

**Figure 4.** Mean Shannon index across all samples per site for
observed and simulated data. Each plot shows one pattern represented by
one exemplary model of each category (overperdicted, intermediate and
underpredicted). The full data with the grouping of the models are shown
in Figures S3 and S4. n indicates the number of models falling in each
group. A) Examples for the three trends across models for species
diversity at the stand level. B) Examples for the three trends across
models for regeneration at the 7 cm threshold.

**Figure 5.** Ratio of tree regeneration rates between the 7 and 10 cm
thresholds. Dashed blue lines mark a ratio equal to 1, indicating no
decrease in tree regeneration between 7 and 10 cm, and a ratio equal to
1.77, corresponding to the Reineke self-thinning ratio under even-aged
conditions.

**Figure 6.** Ratio of tree regeneration between the 7 and 10 cm
thresholds (regeneration 7 cm / regeneration 10 cm) and the
overestimation proportion at 7 cm ([simulated - observed] / observed)
for the mean regeneration per model across sites and samples. The
horizontal dashed lines show a ratio equal to 1, indicating no decrease
of regeneration between 7 and 10 cm, and a ratio equal to 1.77
corresponding to the Reineke self-thinning line under even-aged
conditions. The vertical line indicates zero overestimation.

**Figure 7.** Mean regeneration values across the 200 samples per site
with the y-axis scaled differently by model, for the 200 sites against
gradients of (a) total basal area; (b) climatic water balance; (c)
seasonal degree-day sum. The values were split into ten bins; the red
lines represent a GAM model showing the trend in the observed data. For
comparing the magnitude of recruitment across models the same graphs
with equally scaled y-axis can be found in Figure S7.

**Figure 8.** Regeneration levels of the five main species in terms of
their basal area share in the observations (top row) and the performance
of each model across the environmental gradients (other rows) for the 7
cm regeneration threshold. The values shown are the mean of the 200
samples per site and across the sites in each bin (tile), with ten bins
per gradient. The sizes of the circles represent the ratio between the
regeneration basal area of the species and the total regeneration basal
area for all species. The absence of a circle indicates a zero basal
area share in the regeneration, or the absence of regeneration
altogether. The color gradient (for the models only) shows the
difference between the simulated and observed ratio of regeneration
basal area of the species and the total regeneration basal area for all
the species.

# Figures

```{r map, echo=FALSE, fig.cap=".", out.width='80%'}
knitr::include_graphics(here::here("figures", "theMap.png"))
```

```{r over, echo=FALSE, fig.cap=".", out.width='100%'}
        knitr::include_graphics(here::here("figures", "recOverAll710.png"))
```

```{r H710, echo=FALSE, fig.cap=".", out.width='100%'}
    knitr::include_graphics(here::here("figures", "H7_10.png"))
```

```{r combinedRichness, echo=FALSE, fig.cap=".", out.width='75%'}
knitr::include_graphics(here::here("figures", "combined_richness.png"))
```

```{r mortality11, echo=FALSE, fig.cap=".", out.width='75%'}
knitr::include_graphics(here::here("figures", "ratio7_10.png"))
```

```{r mortality13, echo=FALSE, fig.cap=".", out.width='80%'}
knitr::include_graphics(here::here("figures", "mort7_10_medianOver.png"))
```

```{r nvTrend710BAlightTemp, echo=FALSE, fig.cap=".", fig.show = 'hold', out.width="49%",fig.ncol = 2, fig.subcap=c('', '', ''), fig.show='hold', fig.align='left'}

```{r diff, echo=FALSE, fig.cap=".", out.width='100%', out.height= '85%'}
knitr::include_graphics(here::here("figures", "HeatEnvGrad_10circles_7.png"))
```
